---
jupyter:
  jupytext:
    formats: ipynb,md:myst,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.10.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Setup


## Files

Project base directory:

```{r}
projectdir = '/home/rep/'
```

**Input files:**

```{r}
TACO1              = paste0(projectdir, 'data/original/Study1 Original 22 Experiments/TreatmentAssignment_Covariate_Outcome_Study1.csv')
TACO2              = paste0(projectdir, 'data/original/Study2 11 New Experiments/TreatmentAssignment_Covariate_Outcome_Study2.csv')
model1_predictions = paste0(projectdir, 'data/processed/model_predictions1.csv')
model2_predictions = paste0(projectdir, 'data/processed/model_predictions2.csv')
```

**Output files:**

```{r}
cleaned_data       = paste0(projectdir, 'data/processed/cleaned_data.rda')
```

## Packages

```{r}
library(dplyr)
```

# Merge Data


Here we merge the data from the 22 experiments in Study 1 and the 11 experiments from Study 1, and also add in the predictions obtained from the remnant models.

```{r}
## combine remnant predictions w rest of data for first 22 experiments

### put predictions into main dataset
df <- read.csv(TACO1)
pred <- read.csv(model1_predictions)

pred <- subset(pred,pred$next_is_experiment == 1)
pred$link <- as.character(pred$link)
df$id <- as.character(paste(df$User.ID,df$problem_set,sep='_'))

dat <- df
dat$p_complete <- rep(NA,nrow(dat))

for (i in 1:nrow(df)) {
  p = subset(pred,as.character(pred$link) == df$id[i])
  if (nrow(p) > 1) {
   print(paste('problem with',df$id,sep=' '))
  }
  else if (nrow(p) == 1) {
    dat$p_complete[i] <- p$p_complete[1]
    dat$p_inverted_mastery[i] <- p$p_inv_mastery[1]
  }
  else {
    sub_p <- subset(pred,pred$experiment==df$problem_set[i])
    dat$p_complete[i] <- mean(na.omit(sub_p$complete))
    dat$p_inverted_mastery[i] <- mean(na.omit(sub_p$p_inv_mastery))
  }
}

dat$p_complete <- ifelse(dat$p_complete > 1, 1,dat$p_complete)
dat$p_complete <- ifelse(dat$p_complete < 0, 0,dat$p_complete)
```

```{r}
## combine remnant predictions w rest of data for remaining 11 experiments

pred2 <- read.csv(model2_predictions)
dat2 <- read.csv(TACO2)

dat2 <- merge(pred2,dat2,all.y=TRUE,all.x=FALSE)

dat2$p_complete <- dat2$pcomplete
```

```{r}
dat2$problem_set <- dat2$target_sequence_id

dat$condition <- ifelse(dat$InferredCondition=='E',1,0)

dat <- rbind(dat[,intersect(names(dat),names(dat2))],dat2[,intersect(names(dat),names(dat2))])

## this is the list of experiments
dat$problem_set <- as.factor(dat$problem_set)

### analysis for "complete"; drop subjects for whom "predicted" complete is NA
dat <- droplevels(dat[!is.na(dat$p_complete),])

dat$male <- dat$Guessed.Gender=='\"Male\"'
dat$unknownGender <- dat$Guessed.Gender=='\"Uknown\"'

dat$treatment <- dat$condition

dat$residual <- dat$complete-dat$p_complete
```

# Imputing Missing Covariates


To impute missing covariate values, when possible we imputed the classroom mean covariate value for students working on that skill builder.  When there were no other available values for a covariate for students in the same classroom working on the same skill builder, we imputed with the global mean of students working on that skill builder.  Since covariates are all pre-treatment and the imputation did not depend on treatment status, the imputed covariates are themselves covariates, measured for all subjects.  Therefore, we need not correct for the imputation scheme in our treatment effect estimation.


Names of covariates for within-sample covariate adjustment:

```{r}
covNames <- c(
    "Prior.Problem.Count",
    "Prior.Percent.Correct",
    "Prior.Assignments.Assigned",
    "Prior.Percent.Completion",
    "Prior.Class.Percent.Completion",
    "Prior.Homework.Assigned",
    "Prior.Homework.Percent.Completion",
    "Prior.Class.Homework.Percent.Completion",
    "male",
    "unknownGender")
```

Imputation:

```{r}
dat_preimpute <- dat     # Save original version, too.

### first fill in with class/problem_set mean
### if that doesn't work, fill in with problem_set mean
dat <- dat%>%
  group_by(Class.ID,problem_set)%>%
  mutate(
    across(all_of(covNames),~ifelse(is.finite(.),.,mean(.,na.rm=TRUE)))
  )%>%
  group_by(problem_set)%>%
  mutate(
    across(all_of(covNames),~ifelse(is.finite(.),.,mean(.,na.rm=TRUE)))
  )%>%
  ungroup()

stopifnot(all(sapply(covNames,function(x) mean(is.finite(dat[[x]])))==1))

dat_postimpute <- dat
```

```{r}
save(dat_preimpute, dat_postimpute, covNames, file=cleaned_data)
```
