---
jupyter:
  jupytext:
    formats: ipynb,md:myst,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.10.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Setup


## Files

Project base directory:

```{r}
projectdir = '/home/rep/'
```

**Input files:**

```{r}
cleaned_data = paste0(projectdir, 'data/processed/cleaned_data.rda')
```

**Output files:**

```{r}
fullresfilename = paste0(projectdir, 'output/results/fullres.RData')
olsfilename     = paste0(projectdir, 'output/results/ols.RData')
```

## Packages

```{r}
library(dplyr)
library(loop.estimator)
library(estimatr)
```

## Load data

Load data from R data file:

```{r}
load(cleaned_data)
ls()
```

Use data with imputed missing covariates

```{r}
dat = dat_postimpute
dim(dat)
head(dat,10)
```

# Main Analysis

This is the main analysis, which ultimately is used to produce Figures 1 and 2 in the paper.

## Define Estimators

Here we define the estimators, using a standard format.  Most of these are wrapper functions to functions in the `loop.estimator` package.


$\hat{\tau}^{\textrm{DM}}$ (difference-in-means)

```{r}
simpDiff <- function(datPS,covsPS,outcome='complete'){
  Y <- datPS[[outcome]]
  Z <- datPS$treatment
  est <- mean(Y[Z==1])- mean(Y[Z==0])
  Vhat <- var(Y[Z==1])/sum(Z==1)+var(Y[Z==0])/sum(Z==0)
  return(c(tauhat=est,varhat=Vhat))
}
```

$\hat{\tau}^{\textrm{RE}}$ (difference-in-means on residuals)

```{r}
rebar <- function(datPS,covsPS)
  simpDiff(datPS,covsPS,outcome='residual')
```

$\hat{\tau}^{\textrm{SS}}[x^r, \textrm{LS}]$

```{r}
strat1 <- function(datPS,covsPS)
    with(datPS,
         loop(complete,treatment,as.matrix(p_complete),pred=loop_ols))
```

$\hat{\tau}^{\textrm{SS}}[\tilde{\bf{x}}, \textrm{RF}]$

```{r}
strat2 <- function(datPS,covsPS)
    with(datPS,
         loop(complete,treatment,cbind(p_complete,covsPS)))
```

$\hat{\tau}^{\textrm{SS}}[\tilde{\bf{x}}, \textrm{EN}]$

```{r}
strat3 <- function(datPS,covsPS)
    with(datPS,
         loop(complete,treatment,covsPS,pred=reloop,yhat=p_complete))
```

$\hat{\tau}^{\textrm{SS}}[\mathbf{x}, \textrm{RF}]$

```{r}
justCovs <- function(datPS,covsPS)
    with(datPS,
         loop(complete,treatment,covsPS))
```

## Estimate Effects


Here we estimate effects of treatment for each of the 33 skill builders in the dataset.

First create a function `full()` which computes all five treatment effect estimates discussed in the paper:

```{r}
## all the estimators, for a particular problem set
full <- function(ps,dat,covNames,methods=c('simpDiff','rebar','strat1','strat3','justCovs')){
    if(!'simpDiff'%in%methods) methods=c('simpDiff',methods)

    datPS <- dat[dat$problem_set==ps,]
    covsPS <- dat[dat$problem_set==ps,covNames]

    res <- sapply(methods,
                  function(FUN){
                      set.seed(613)
                      fun <- get(FUN)
                      out <- try(fun(datPS,covsPS))
                      if(inherits(out,'try-error')) return(rep(NA,2))
                      out
                  },simplify=FALSE)
    res <- do.call('rbind',res)
    res[,2] <- sqrt(res[,2])  ## standard error, not variance
    colnames(res) <- c('est','se')
    res <- cbind(res,improvement=1-res[,'se']/res['simpDiff','se'])
    attr(res,'psid') <- ps
    attr(res,'n') <- nrow(datPS)
    res
}
```

Now compute the estimates:

```{r}
fullres <- sapply(levels(dat$problem_set),full,dat=dat,
                  covNames=covNames,simplify=FALSE)

### name the problem sets based on the 
### SE from the simple difference estimator
rnk <- rank(sapply(fullres,function(x) x['simpDiff','se']))
names(fullres) <- as.character(rnk)

for(i in 1:length(fullres))
    attr(fullres[[i]],'psid') <- levels(dat$problem_set)[i]

dat$ps <- rnk[as.character(dat$problem_set)]
    
dat_withps <- dat
```

Save results

```{r}
save(fullres, dat_withps, file=fullresfilename)
```

# Comparing Sample Splitting to ANCOVA Estimators


The following is to create Figure 3 in Section 4.3.

## Define Estimators

First we need to define some additional estimators.  These are wrappers to the functions in the `estimatr` package.

```{r}
reloopLin <- function(datPS,covsPS){
    est <- lm_lin(complete~treatment,covariates=~p_complete,data=datPS)
    c(est=est$coefficients['treatment'],var=est$std.error['treatment']^2)
}

reloopAlt <- function(datPS,covsPS){
    est <- lm_robust(complete~treatment+p_complete,data=datPS)
    c(est=est$coefficients['treatment'],var=est$std.error['treatment']^2)
}

reloopPlusLin <- function(datPS,covsPS){
    est <- lm_lin(complete~treatment,covariates=~p_complete+as.matrix(covsPS),data=datPS)
    c(est=est$coefficients['treatment'],var=est$std.error['treatment']^2)
}

reloopPlusAlt <- function(datPS,covsPS){
    est <- lm_robust(complete~treatment+p_complete+as.matrix(covsPS),data=datPS)
    c(est=est$coefficients['treatment'],var=est$std.error['treatment']^2)
}

lin <- function(datPS,covsPS){
    est <- lm_lin(complete~treatment,covariates=~as.matrix(covsPS),data=datPS)
    c(est=est$coefficients['treatment'],var=est$std.error['treatment']^2)
}

ancova <- function(datPS,covsPS){
    est <- lm_robust(complete~treatment+as.matrix(covsPS),data=datPS)
    c(est=est$coefficients['treatment'],var=est$std.error['treatment']^2)
}
```

## Estimate Effects

This estimates the effects and their SEs:

```{r}
dat = dat_postimpute

dat <- dat%>%
  group_by(Class.ID,problem_set)%>%
  mutate(
    across(all_of(covNames),~ifelse(is.finite(.),.,mean(.,na.rm=TRUE)))
  )%>%
  group_by(problem_set)%>%
  mutate(
    across(all_of(covNames),~ifelse(is.finite(.),.,mean(.,na.rm=TRUE)))
  )%>%
  ungroup()



ols <- sapply(levels(dat$problem_set),full,dat=dat,
covNames=covNames,simplify=FALSE,
                  methods=c('reloopLin','reloopAlt','reloopPlusLin',
                  'reloopPlusAlt','lin','ancova'))
```

Save results

```{r}
save(ols,file=olsfilename)
```
