---
output: 
  pdf_document:
    extra_dependencies:
      tikz: null
      bm: null
title: Tables, Figures, and Numerical results
jupyter:
  jupytext:
    formats: ipynb,md:myst,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.10.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Setup

## Files

Project base directory:

```{r}
projectdir = '/home/rep/'
```

**Input files:**

```{r}
cleaned_data    = paste0(projectdir, 'data/processed/cleaned_data.rda')
fullresfilename = paste0(projectdir, 'output/results/fullres.RData')
olsfilename     = paste0(projectdir, 'output/results/ols.RData')
```

**Output directories:**

```{r}
resultsdir = paste0(projectdir, 'output/results/')
tabledir   = paste0(projectdir, 'output/tables/')
figuredir  = paste0(projectdir, 'output/figures/')
```

Latex notation file.  Used when producing tables and `tikz` figures.

```{r}
latexnotation = paste0(projectdir, 'misc/notation.tex')
```

\input{../misc/notation.tex}

## Packages

```{r}
library(scales)
library(dplyr)
library(ggplot2)
library(tibble)
library(purrr)
library(tidyr)
library(kableExtra)
library(xtable)
library(IRdisplay)
library(tikzDevice)
library(forcats)
```

## Load data

Load data from R data files:

```{r}
load(cleaned_data)
load(fullresfilename)
load(olsfilename)
ls()
```

# Numerical Results

The following code creates a dataset called `comparisons` that includes the sampling variance ratios comparing each method to the others, for each problem set.  It also produces a table (which is not in the manuscript) giving the estimated standard error for each method and each experiment.

```{r}
methodName=c(
               ReLOOPEN='\\trcpen',
               ReLOOP='\\trc',
               Rebar='\\trebar',
               Loop='\\tss[\\bx,\\mathrm{RF}]',
               SimpleDifference='\\tsd'
           )
           
pd <- map_dfr(
    names(fullres),
    function(x) cbind(as.data.frame(fullres[[x]]),
      method=factor(rownames(fullres[[x]]),
        levels=c('simpDiff','justCovs','rebar','strat1','strat2','strat3')),
      rnk=x))%>%
    select(-improvement)%>%
    rename(experiment=rnk)%>%
    filter(method!='strat2')%>%
    mutate(method=fct_recode(.$method,`ReLOOP`='strat1',`ReLOOPEN`='strat3',Loop='justCovs',Rebar='rebar',`SimpleDifference`='simpDiff'))%>%
    mutate(method=factor(method,c( "ReLOOPEN", "ReLOOP", "Loop", "Rebar", "SimpleDifference" ), ordered = TRUE )
           )

pd3 <- pd %>% 
  group_by(experiment)%>%
  mutate(ssMult=se[method=='SimpleDifference']^2/se^2,
         percent = se / se[method=='SimpleDifference'] ) %>%
  ungroup() %>%
  group_by(method) %>%
  mutate(lab=as.character(experiment)) %>%
  ungroup()

pwide = pd3 %>% dplyr::select(method, experiment, se ) %>% spread( method, se )

save(pwide,file=paste0(resultsdir, 'SEs.RData'))

make_comp = function( A, B ,pwide) {
    data.frame(comp = paste0('\n$\\frac{\\varhat(',methodName[as.character(B)],')}{\\varhat(',methodName[as.character(A)],')}$\n'),
               experiment=pwide$experiment,
               ssMult = pwide[[B]]^2 / pwide[[A]]^2,
               method1=as.character(A),
               method2=as.character(B),
               stringsAsFactors = FALSE )
}

combos = expand.grid( A = levels( pd3$method ),
                      B = levels( pd3$method ) )
combos = filter( combos, as.numeric(A) < as.numeric(B) )
combos = mutate( combos,
                 A = as.character(A),
                 B = as.character(B) )

comparisons = pmap_df( combos, make_comp ,pwide=pwide)

## get the right order
methodOrd=rev(c('SimpleDifference','Rebar','Loop','ReLOOP','ReLOOPEN'))
comparisons$method1 <- factor(comparisons$method1,levels=methodOrd)
comparisons$method2 <- factor(comparisons$method2,levels=methodOrd)

compLevs=rev(unique(comparisons$comp[order(as.numeric(comparisons$method2),as.numeric(comparisons$method1))]))

comparisons$comp <- factor(comparisons$comp,levels=compLevs)

save(comparisons,file=paste0(resultsdir, 'reloopComparisons.RData'))
```

The following code reproduces some of the numbers in the manuscript text describing the results:

```{r}
compTab <- comparisons %>% 
  group_by(method1,method2) %>%
  summarize(
            worse=sum(ssMult<0.975),
            equal=sum(abs(ssMult-1)<0.025),
            better=sum(ssMult>1.025),
            best=max(ssMult),
            bestPS=experiment[which.max(ssMult)],
            best2=sort(ssMult,decreasing=TRUE)[2],
            best2ps=experiment[rank(ssMult)==32],
            worst=min(ssMult),
            worstPS=experiment[which.min(ssMult)]
            ) %>%
    ungroup() %>%
    mutate(across(starts_with('method'),
        ~paste0('$',methodName[as.character(.)],'$')))

compTab %>% 
    select(method1:bestPS) %>% 
    kable(escape = FALSE, format="latex") %>% 
    as.character %>% 
    cat(file=paste0(tabledir, 'compTab1.tex'))
    
compTab %>% 
    select(method1,method2,best2:worstPS) %>% 
    kable(escape=FALSE, format="latex") %>% 
    as.character %>% 
    cat(file=paste0(tabledir, 'compTab2.tex'))
```


```{r}
pwidePrint <- pwide
names(pwidePrint)[-1] <- paste0('$',methodName[names(pwidePrint)[-1]],'$')
SEtab = kable(pwidePrint,row.names=FALSE,
     caption="Estimated standard error for the ATE
     in each skill builder, using each method 
     discussed in the manuscript",
     label="tab:SEs",digits=3,escape=FALSE,format="latex")
cat(SEtab, file=paste0(tabledir, 'SETab.tex'))
```

Display the results here (if compiling to latex):

\input{../output/tables/compTab1.tex}

\input{../output/tables/compTab2.tex}

\input{../output/tables/SETab.tex}


# Tables


### Table 1

```{r}
dat = dat_preimpute

covTable <-
    covNames%>%
    setdiff(c('male','unknownGender'))%>%
    map_dfr(
        .,
        function(x){
            tibble(
                Covariate=gsub('Prior.','',x,fixed=TRUE)%>%gsub('.',' ',.,fixed=TRUE),
                Mean=mean(dat[[x]],na.rm=TRUE),
                SD=sd(dat[[x]],na.rm=TRUE),
                `% Missing`=round(mean(is.na(dat[[x]]))*100)
                )
        }
    )%>%
    column_to_rownames("Covariate")%>%
    xtable(
        digits=c(0,2,2,0),
        caption="Summary statistics for aggregate prior ASSISTments performance used as within-sample covariates: number of problems worked, and assignments and homework assigned, percent of problems correct on first try, assignments completed, and homework completed at the student and class level, and students' genders, as guessed by ASSISTments based on first names.",
        label='tab:covariates'
    )

ATR <- list(
              pos=list(8),
              command=paste0(
                  'Guessed Gender',
                  '&Male: ', round(mean(dat$Guessed.Gender=='\"Male\"')*100),'\\%',
                  '&Female: ',round(mean(dat$Guessed.Gender=='\"Female\"')*100),'\\%',
                  '&Unknown: ',round(mean(dat$Guessed.Gender=='\"Uknown\"')*100),'\\%',
                  '\\\\\n'
              )
    )

print(covTable, add.to.row=ATR, file=paste0(tabledir, 'covtab.tex'))
```

\input{../output/tables/covtab.tex}

### Table 2

```{r}
dat = dat_withps #The numbering of the experiments derives from the estimated standard errors, so use dat_withps, which comes after effect estimation.

tab <- dat%>%
    group_by(ps,treatment)%>%
    summarize(n=n(),percomplete=round(mean(complete)*100))%>%
    ungroup()%>%
    pivot_wider(names_from=treatment,values_from=c(n,percomplete))

tab <- cbind(tab[1:17,],rbind(tab[18:33,],rep(NA,5)))

ttt <- kbl(tab,
           format='latex',
           booktabs=TRUE,
           col.names=rep(c("Experiment",rep(c("Trt","Ctl"),2)),2),
           caption="Sample sizes and \\% homework completion---the outcome of interest---by treatment group in each of the 33 A/B tests.",          
           label="info")%>%
    kable_styling()%>%
    column_spec(5,border_right=TRUE)%>%
    add_header_above(rep(c(" "=1,"n"=2,"% Complete"=2),2))

cat(ttt, file=paste0(tabledir, 'infoTab.tex'))
```

\input{../output/tables/infoTab.tex}

# Figures

```{r}
options(list(tikzLatexPackages=c(
  getOption('tikzLatexPackages'),
  '\\usepackage{amsmath,amsfonts,amsthm,amssymb,thmtools}',
  '\\usepackage{bm}', 
  readLines(latexnotation))))
```

### Figure 1

```{r}
p <- comparisons%>%
    filter(method1%in%c('ReLOOP','Rebar'),
    method2%in%c('ReLOOPEN','Rebar','SimpleDifference'))%>%
    ggplot(aes(ssMult))+
    geom_dotplot( method="histodot", binwidth = .05 )  +
    labs( x = "Relative Ratio of Sample Variances", y="" ) +
    geom_vline( xintercept = 1, col="red" ) +
    facet_wrap(~comp,nrow=1)+
    theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y= element_blank(),
        axis.ticks.y = element_blank(),
        text=element_text(size=12),
        strip.text=element_text(size=12,lineheight=0.5))

tikz(paste0(figuredir, 'fig4.tex'),width=6.4,height=2,standAlone=FALSE)
print(p)
dev.off()
```

\input{../output/figures/fig4.tex}

### Figure 2

```{r}
p <- comparisons%>%
    filter(method1%in%c('ReLOOPEN'),
    method2%in%c('Loop','ReLOOP','SimpleDifference'))%>%
    mutate(comp=factor(comp,levels=unique(as.character(comp))))%>%
ggplot(aes(ssMult))+
    geom_dotplot( method="histodot", binwidth = .05 )  +
    labs( x = "Relative Ratio of Sample Variances", y="" ) +
    geom_vline( xintercept = 1, col="red" ) +
    facet_wrap(~comp,nrow=1)+
    theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y= element_blank(),
        axis.ticks.y = element_blank(),
        text=element_text(size=12),
        strip.text=element_text(size=12,lineheight=0.5))

tikz(paste0(figuredir, 'fig5alt.tex'),width=6.4,height=2,standAlone=FALSE)
print(p)
dev.off()
```

\input{../output/figures/fig5alt.tex}

### Figure 3


The following creates Figure 3 (plus some other plots)

This estimates the effects and their SEs:

```{r}
methodName=c(
               ReLOOPEN='\\hat{\\tau}^{\\mathrm{SS}}[\\tilde{\\bm{x}},\\mathrm{EN}]',
               ReLOOP='\\hat{\\tau}^{\\mathrm{SS}}[x^r,\\mathrm{LS}]',
               Rebar='\\trebar',
               Loop='\\hat{\\tau}^{\\mathrm{SS}}[\\bm{x},\\mathrm{RF}]',
               SimpleDifference='\\hat{\\tau}^{\\mathrm{DM}}'
           )

methodNameLin=c(
               ReLOOPEN='\\tau^{Lin}[\\tilde{\\bm{x}}]',
               ReLOOP='\\tau^{Lin}[x^r]',
               Loop='\\tau^{Lin}[\\bm{x}]',
               SimpleDifference='\\hat{\\tau}^{\\mathrm{DM}}'
           )

methodNameAlt=c(
               ReLOOPEN='\\hat{\\beta}[\\tilde{\\bm{x}}]',
               ReLOOP='\\hat{\\beta}[x^r]',
               Loop='\\hat{\\beta}[\\bm{x}]',
               SimpleDifference='\\hat{\\tau}^{\\mathrm{DM}}'
           )

lin <- sapply(ols,
              function(x){
                x <- x[c('simpDiff','reloopLin','reloopPlusLin','lin'),]
                rownames(x) <- c('simpDiff','strat1','strat3','justCovs')
                x},
              simplify=FALSE)

alt <- sapply(ols,
              function(x){
                x <- x[c('simpDiff','reloopAlt','reloopPlusAlt','ancova'),]
                rownames(x) <- c('simpDiff','strat1','strat3','justCovs')
                x},
              simplify=FALSE)


pdLin <- map_dfr(names(lin),
    function(x) cbind(as.data.frame(lin[[x]]),
                      method=factor(rownames(lin[[x]]),
        levels=c('simpDiff','justCovs','rebar','strat1','strat2','strat3')),
      rnk=x))%>%
    select(-improvement)%>%
    rename(experiment=rnk)%>%
    filter(method!='strat2')%>%
    mutate(method=fct_recode(.$method,`ReLOOP`='strat1',`ReLOOPEN`='strat3',Loop='justCovs',Rebar='rebar',`SimpleDifference`='simpDiff'))%>%
    mutate(method=factor(method,c( "ReLOOPEN", "ReLOOP", "Loop", "Rebar", "SimpleDifference" ), ordered = TRUE )
           )

pdAlt <- map_dfr(names(alt),
    function(x) cbind(as.data.frame(alt[[x]]),
                      method=factor(rownames(alt[[x]]),
        levels=c('simpDiff','justCovs','rebar','strat1','strat2','strat3')),
      rnk=x))%>%
    select(-improvement)%>%
    rename(experiment=rnk)%>%
    filter(method!='strat2')%>%
    mutate(method=fct_recode(.$method,`ReLOOP`='strat1',`ReLOOPEN`='strat3',Loop='justCovs',Rebar='rebar',`SimpleDifference`='simpDiff'))%>%
    mutate(method=factor(method,c( "ReLOOPEN", "ReLOOP", "Loop", "Rebar", "SimpleDifference" ), ordered = TRUE )
           )


pd3Lin <- pdLin%>%group_by(experiment)%>%
  mutate(ssMult=se[method=='SimpleDifference']^2/se^2,
         percent = se / se[method=='SimpleDifference'] ) %>%
  ungroup() %>%
  group_by(method) %>%
  mutate(lab=as.character(experiment)) %>%
  ungroup()

pd3Alt <- pdAlt%>%group_by(experiment)%>%
  mutate(ssMult=se[method=='SimpleDifference']^2/se^2,
         percent = se / se[method=='SimpleDifference'] ) %>%
  ungroup() %>%
  group_by(method) %>%
  mutate(lab=as.character(experiment)) %>%
  ungroup()

pwideLin = pd3Lin %>% dplyr::select(method, experiment, se ) %>%
  pivot_wider('experiment',names_from='method',values_from='se')

pwideAlt = pd3Alt %>% dplyr::select(method, experiment, se ) %>%
  pivot_wider('experiment',names_from='method',values_from='se')


make_comp = function( A, B ,pwide,methodName) {
    data.frame( comp = paste0('\n$\\frac{\\hat{\\mathbb{V}}(',methodName[as.character(B)],')}{\\hat{\\mathbb{V}}(',methodName[as.character(A)],')}$\n'),
             experiment=pwide$experiment,
             ssMult = pwide[[B]]^2 / pwide[[A]]^2,
             method1=as.character(A),
             method2=as.character(B),
             stringsAsFactors = FALSE )
}

combos = expand.grid( A = levels( droplevels(pd3Lin$method )),
                      B = levels( droplevels(pd3Lin$method ) ))
combos = filter( combos, as.numeric(A) < as.numeric(B) )
combos = mutate( combos,
                 A = as.character(A),
                 B = as.character(B) )

comparisonsLin = pmap_dfr( combos, make_comp ,pwide=pwideLin,methodName=methodNameLin)
comparisonsAlt = pmap_dfr( combos, make_comp ,pwide=pwideAlt,methodName=methodNameAlt)

### get the right order
methodOrd=rev(c('SimpleDifference','Rebar','Loop','ReLOOP','ReLOOPEN'))
comparisonsLin$method1 <- factor(comparisonsLin$method1,levels=methodOrd)
comparisonsLin$method2 <- factor(comparisonsLin$method2,levels=methodOrd)

compLevs=rev(unique(comparisonsLin$comp[order(as.numeric(comparisonsLin$method2),as.numeric(comparisonsLin$method1))]))

comparisonsLin$comp <- factor(comparisonsLin$comp,levels=compLevs)


comparisonsAlt$method1 <- factor(comparisonsAlt$method1,levels=methodOrd)
comparisonsAlt$method2 <- factor(comparisonsAlt$method2,levels=methodOrd)

compLevs=rev(unique(comparisonsAlt$comp[order(as.numeric(comparisonsAlt$method2),as.numeric(comparisonsAlt$method1))]))

comparisonsAlt$comp <- factor(comparisonsAlt$comp,levels=compLevs)

### compare to reloop

crosswalk <- sapply(fullres,function(x) attributes(x)$psid)
pwideLin$experiment=vapply(pwideLin$experiment,function(ee) names(crosswalk)[crosswalk==ee],'a')

pwideCompLin <- merge(pwide,pwideLin,by='experiment',suffixes=c('','Lin'))
comps <- data.frame(A=names(pwideLin)[-c(1,2)],B=paste0(names(pwideLin)[-c(1,2)],'Lin'))

namesReloopLin <- c(methodName,setNames(methodNameLin,paste0(names(methodNameLin),'Lin')))

comparisonsReloopLin <- pmap_dfr(comps,make_comp,pwide=pwideCompLin,methodName=namesReloopLin)

pwideAlt$experiment=vapply(pwideAlt$experiment,function(ee) names(crosswalk)[crosswalk==ee],'a')

pwideCompAlt <- merge(pwide,pwideAlt,by='experiment',suffixes=c('','Alt'))
comps <- data.frame(A=names(pwideLin)[-c(1,2)],B=paste0(names(pwideLin)[-c(1,2)],'Alt'))

namesReloopAlt <- c(methodName,setNames(methodNameAlt,paste0(names(methodNameLin),'Alt')))

comparisonsReloopAlt <- pmap_dfr(comps,make_comp,pwide=pwideCompAlt,methodName=namesReloopAlt)


newcomp <- bind_rows(
  filter(comparisons,method2=='SimpleDifference',startsWith(as.character(method1),'ReLOOP')),
  filter(comparisonsAlt,method2=='SimpleDifference',startsWith(as.character(method1),'ReLOOP')),
  filter(comparisonsReloopAlt,startsWith(as.character(method1),'ReLOOP')))


newcomp$comp <- gsub('\\\\varhat','\\\\hat{\\\\mathbb{V}}',newcomp$comp)
newcomp$comp <- gsub('\\\\trcpen','\\\\hat{\\\\tau}^{\\\\mathrm{SS}}[\\\\tilde{\\\\bm{x}},\\\\mathrm{EN}]',newcomp$comp)
newcomp$comp <- gsub('\\\\trc','\\\\hat{\\\\tau}^{\\\\mathrm{SS}}[x^r,\\\\mathrm{LS}]',newcomp$comp)
newcomp$comp <- gsub('\\\\tsd','\\\\hat{\\\\tau}^{\\\\mathrm{DM}}',newcomp$comp)

newcomp$comp <- factor(newcomp$comp,
                       levels=c(
                         '\n$\\frac{\\hat{\\mathbb{V}}(\\hat{\\tau}^{\\mathrm{DM}})}{\\hat{\\mathbb{V}}(\\hat{\\tau}^{\\mathrm{SS}}[x^r,\\mathrm{LS}])}$\n',
                         '\n$\\frac{\\hat{\\mathbb{V}}(\\hat{\\tau}^{\\mathrm{DM}})}{\\hat{\\mathbb{V}}(\\hat{\\tau}^{\\mathrm{SS}}[\\tilde{\\bm{x}},\\mathrm{EN}])}$\n',
                         '\n$\\frac{\\hat{\\mathbb{V}}(\\hat{\\tau}^{\\mathrm{DM}})}{\\hat{\\mathbb{V}}(\\hat{\\beta}[x^r])}$\n',
                         '\n$\\frac{\\hat{\\mathbb{V}}(\\hat{\\tau}^{\\mathrm{DM}})}{\\hat{\\mathbb{V}}(\\hat{\\beta}[\\tilde{\\bm{x}}])}$\n',
                         '\n$\\frac{\\hat{\\mathbb{V}}(\\hat{\\beta}[x^r])}{\\hat{\\mathbb{V}}(\\hat{\\tau}^{\\mathrm{SS}}[x^r,\\mathrm{LS}])}$\n',
                         '\n$\\frac{\\hat{\\mathbb{V}}(\\hat{\\beta}[\\tilde{\\bm{x}}])}{\\hat{\\mathbb{V}}(\\hat{\\tau}^{\\mathrm{SS}}[\\tilde{\\bm{x}},\\mathrm{EN}])}$\n'))
```

```{r}
pOls3 <- ggplot(newcomp,aes(ssMult))+
    geom_dotplot( method="histodot", binwidth = .01 )  +
    labs( x = "Relative Ratio of Sample Variances", y="" ) +
    geom_vline( xintercept = 1, col="red" ) +
    facet_wrap(~comp,nrow=3)+
    theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y= element_blank(),
        axis.ticks.y = element_blank(),
        text=element_text(size=12),
        strip.text=element_text(size=12,lineheight=0.5))+
  scale_x_continuous(trans="log10",breaks=c(0.85,1,1.2,1.4,1.7,2,2.5))

tikz(paste0(figuredir, 'OlsReloop.tex'),width=5,height=6,standAlone=FALSE)
print(pOls3)
dev.off()
```

\input{../output/figures/OlsReloop.tex}
